---
title: "K-12 shootings"
author: "Captain Peaches"
date: "December 11, 2018"
output: html_document
---

## Setup & Data

```{r setup-1, echo=FALSE}
library(tidyverse)
library(DescTools)
library(DataExplorer)
```

Load data

```{r data, echo=FALSE}
dat <- read.csv("K-12 shootings (Public) - K-12 SSDB (Public) Linked.csv", sep = ",")
```

# Take a peek

First exmaine the structure of the data, as well as the type and structure of the variables. 
```{r}
str(dat)
```

I see many variables we might be interested in. Initially, I'm curious to know the distribution of the reasons the shooting occured. 

First, lets check our dataset for missing parameters and cleanliness. 

```{r clean}
plot_intro(dat)
```

Lets also check our variables for correlation, this will make it easier to confirm if a question is possible to answer or not. 

 

```{r correlation}
dat %>%
  plot_correlation(type = "c", maxcat = 5)
```

** Need to fix this plot 

# Question 1

What is the distribution for reasons the shooting was committed? is there an outlying reason shootings occur seemingly more than anywhere else in the world? 

What is the preferred type of weapon/firearm used by the shooters in our ds?

```{r}
dat %>%
  ggplot(aes(x = Firearm.Type)) + 
  geom_histogram(stat = "count") +
  labs(title = "Type of Firearms used in K-12 Incidents", x = "Firearm type", y = "count") + scale_x_discrete(labels = abbreviate, limits = c("Combination", "Handgun", "Other", "Rifle", "Shotgun"))
```

Handguns were used in easily the majority of these shootings. 

# Question 2

What kind of shootings have the most victims? Reasoning being somewhat stoic, but when shooters are more "angry", does this have any affect on the number of lives they take? 

```{r reason~victims}
dat %>%
  ggplot(aes(x = reorder(as.factor(Category), Total.Injured.Killed.Victims), y = Total.Injured.Killed.Victims)) +
  geom_boxplot(alpha = .3, color = "red") +
  coord_flip() + 
  labs(title = "Number of victims described by incident type", y = "# of Victims", 
         x = "Reason shooting occured") 
```

It seems as though this can be somewhat confirmed, however plausible at best. The main observation is that shootings that occured due to Race, Mental Health, Murder/suicide, Hostages, and illegal drug related tend to have more victims. 

What is most surprising is that shootings deemed "indiscriminate" tend to have the most victims. Could this be a reporting error, as we don't have techniques, etc to recognize the cause behind these shootings? Or is it more likely when the "indiscriminate" shooters are searching out targets, everyone becomes a target. 

Lucky for us, we can check to see in how may of these cases the shooter was targetting anyone specifically. 

```{r}
dat %>%
  ggplot(aes(x = Targeted.Specific.Victim.s., fill = Targeted.Specific.Victim.s.)) +
  geom_bar() + coord_flip() + xlim("Y", "N", "Unkown", "Officer Involved")
```

This seems to contradict our thoughts above, concerning who is the target when the shooting is "indiscriminate". This points to make our "anger hypothesis" more plausible than the former. 















